biological data preperation

### 2018-09-18 src/biological_data_prep.Rmd
unique cell id extracted

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
Th
```{r pre-install packages, message=FALSE}
packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp",)
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

@@

# createa unique cell ID

>>>>

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
Th
```{r pre-install packages, message=FALSE}
packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster", "tibble")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

@@

# create a unique cell ID

A unique cell ID layer has been created in the environmental_data_preperation.Rmd. Now you want to extract the value of that layer to the .csv you are creating...

Load the env. data
```{r}
library(raster)
cell_layer <- raster("../output/env/unique_cell.tif") #this is loading the cell layer that was created and saved as a tif
plot(cell_layer)
```
and just grab the attributes of the cell_layer
```{r}
cell_layer
```

The occurrence data needs to be converted into spatialpoints dataframe
```{r}
library(sp)
xy <- adddata[ ,c(3,2)] #3 = decimalLongitude, 2 = decimalLatitude
adddata_aea <- SpatialPointsDataFrame(coords = xy, data = adddata, proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))
adddata_aea
```

```{r}
adddata_aea <- spTransform(adddata_aea, CRS = "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs") #note -60 used to be -91. Changed to 'straighten up'
adddata_aea
```

and plot and Write the transformed species data to a new csv (this will also have the coordinates in meters - note meters are under decimalLongitude1 and decimalLatitude1)...
```{r}
plot(adddata_aea, axes=TRUE, cex.axis=.95)
write.csv(adddata_aea, file = "../output/bio/data_aea.csv", row.names = FALSE)
```

Ok now plot the env. and point data...
```{r}
plot(cell_layer)
points(adddata_aea$decimalLongitude, adddata_aea$decimalLatitude)
```

ok now the next step is to extract the unique cell value and add to each of the points
then write a .csv
then convert back to a normal dataframe
```{r}
adddata_aea$cell_id <- extract(x=cell_layer, y = adddata_aea)
write.csv(adddata_aea, file = "../output/bio/data_aea_cell.csv", row.names = FALSE)
data_aea <- as.data.frame(adddata_aea)
head(data_aea)
```
looking at the header data, there seems to be some new columns created on the way...
```{r}
colnames(data_aea)
```
I don't want "optional" - get rid of it!
```{r}
data_aea <- subset(data_aea, select = -c(optional))
colnames(data_aea)
write.csv(adddata_aea, file = "../output/bio/data_aea_cell.csv", row.names = FALSE)
```
